---
title: "Clustering"
author: "vidhu"
date: "June 6, 2018"
output: html_document
---



Regression Model
```{r}
Adult_del = Adult[which(complete.cases(Adult)==TRUE),]
model = glm(income~.,data = Adult_del, family= "binomial")
Null =  glm(income~1,data = Adult_del, family= "binomial")
step <- step(Null, scope=list(lower=Null, upper=model), direction= "both", trace=TRUE)

vif(step)
model_1= glm(income~age+workclass+education+occupation+relationship+race+sex+nativecountry+capitalgain_cat+capitalloss_cat+hoursperweek_cat, data = Adult_del, family = "binomial")
model_2 = glm(income~age+workclass+education+maritalstatus+occupation+race+sex+nativecountry+capitalgain_cat+capitalloss_cat+hoursperweek_cat, data = Adult_del, family = "binomial")
model_3 = glm(income~age+workclass+education+occupation+race+sex+nativecountry+capitalgain_cat+capitalloss_cat+hoursperweek_cat, data = Adult_del, family = "binomial")
anova(model, model_3, test = "Chisq")
anova(model,model_2, test = "Chisq")
anova(model, model_1, test = "Chisq")
summ = summary(step)
# log odds and their confidence interval
Y = exp(confint.default(step))
X=exp(coef(step))
XY = data.frame(odds = X, confInt_lower = Y[,1], confInt_upper = Y[,2])
XY$odds = round(XY$odds,4)
XY$confInt_lower = round(XY$confInt_lower,4)
XY$confInt_upper = round(XY$confInt_upper,4)
```

1) Is there any gender inequality in the society, that is, do the females earn less /more than males while working in same occupation and same work class?

2) There is a common belief that education is the key to become richer. Does the people who are highly educated more likely to earn above 50K as compared to others?


3)	Find the people who would likely to make the investment.
```{r}
investment = adult[which(complete.cases(adult)==TRUE),]
investment$investor= 0
investment$investor = ifelse(investment$capitalgain == 0 & investment$capitalloss == 0, 0, 1)
head(investment)
investment$investor = as.factor(investment$investor)

```

```{r}
table(investment$investor)

```

```{r}

mylogitnull = glm(investor~1, data= investment, family = binomial)
summary(mylogitnull)
```

```{r}

mylogitfull = glm(investor~age + workclass+ fnlwgt +educationnum+ maritalstatus +occupation +relationship+ race +sex +hoursperweek_cat + nativecountry +income, investment, family = binomial)
summary(mylogitfull)
anova(mylogitfull, test = "LRT")
```


```{r}
#Removing fnlwgt, race and native country
mylogit1 = glm(investor~age + workclass +educationnum +maritalstatus+ occupation +relationship +sex +hoursperweek + income , investment, family = binomial, x=TRUE)
summary(mylogit1)
anova(mylogitfull, mylogit1, test = "Chisq")
anova(mylogit1, test = "Chisq")
```
p value is not <0.05, therefore models are not different. That means race , native country and fnlwt does not contribute much to the model.


```{r}

mylogit2 <- glm(investor ~ age + education + occupation + relationship +sex, data = investment, family = "binomial")
summary(mylogit2)
```


```{r}

mylogit3 <- glm(investor ~ age +education+ relationship+sex +income, data = investment, family = "binomial")
summary(mylogit3)
anova(mylogit1, mylogit3,test = "Chisq")
```

```{r}

mylogit4 <- glm(investor ~ age +education+ sex +income, data = investment, family = "binomial")
summary(mylogit4)
```
